---
title: "Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
---


```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}

listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel")
paket(listOfPackages)
```

# R session information
```{r}
sessionInfo()
```


```{r}
# options(repr.plot.width = 18, repr.plot.height = 9)
theme_plots <- theme_bw() +
  theme(strip.text = element_text(size = 5),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 10),
        plot.subtitle = element_text(size = 9, face = "italic")) 
theme_set(theme_plots)

# Colorblind palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```


# Inclusion criteria
The inclusion criteria for this study is:
- Patients inclusion criteria: 
  - Age between 11-17 at the date of visit
- Visits inclusion criteria:
  - Inpatients (at least one hospitalization > 1 day)
  - Starting date between January 1st, 2019 and July 31st, 2021
  - Comprising least one ICD-10 code pertaining to the specified ICD-10 codes list (mental health related codes)

# Variables that need to be checked/modified by each site 
Change the values of the following variables according to the specificities of your site:
- "folder_4ce_files": folder path where your phase 2.2 data files are located
- "obfuscation": determine the obfuscation threshold (FALSE if no obfuscation, or the numeric value of the obfuscation threshold if any)
- "raceAvailable": set as TRUE or FALSE depending on whether the variable is being collected at your site
- "dateFormat": specify the format of the date at your site (e.g., for "03-AUG-20", the format would be "%d-%b-%y", [see documentation](https://www.stat.berkeley.edu/~s133/dates.html))

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "/Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
```


```{r}
# cut-off dates used in the analysis, these should not be changed
history_start_date <- "2019-01-01"
history_end_date <- "2021-05-31"
pandemic_start_date <- "2020-03-15"
```

# Data input
We will use as input the 2.2 data files. Specifically: 
- LocalPatientSummary
- LocalPatientObservation
- LocalPatientClinicalcourse

For sites recording the race, we will also use an additional file:
- LocalPatientRace

## Read csv files

### 4CE phase 2.2 files

```{r message=FALSE, warning=FALSE}
### Read the CSV input files
source("R/readInputFiles.R")


files <- readInputFiles( path      = "../Input/09212021_2/", 
                         separator = ",",
                         skip      = 0, 
                         verbose   = FALSE )
  
## Create the output folder if it doesn't exist
#if (! "output" %in% list.dirs()) dir.create("output")

### Extract the patient summary and observation information. 
demo_raw <- files[["patientSummary"]]
obs_raw <- files[["patientObservations"]]
clinical_raw <- files[["patientClinicalCourse"]]

### Read the file containing race information for those sites recording this variable
if( raceAvailable == TRUE ){
  race_raw <- read.delim(paste0("../", folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
}
```


```{r}
# Adding filtering on in_hospital == 1 here to reduce subsequent compute time
clinical_raw <- filter(clinical_raw, in_hospital == 1)
```

### Adding columns with hospitalization number and hospitalization length to the clinical raw file
```{r}
count_sequences_hospitalisation <- function(df, ...) {
  seq_hospitalisation_df <- data.frame(total_span = seq(min(df$days_since_admission),
                                                        max(df$days_since_admission))
  ) %>%
    left_join(df, by = c("total_span" = "days_since_admission")) %>%
    replace_na(list(in_hospital = 0))
  count_sequences <- rle(seq_hospitalisation_df$in_hospital)
  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
  n_sequences <- seq_along(count_sequences_1$lengths)
  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
  stopifnot(length(df$days_since_admission) == length(sequences))
  data.frame(days_since_admission = df$days_since_admission,
             n_hospitalisation = sequences,
             len_hospitalisation = sequences_len)
}
stopifnot(all(clinical_raw$in_hospital == 1))
hospitalisations_seq_df <- clinical_raw %>%
  distinct(patient_num, cohort, days_since_admission, in_hospital) %>%
  group_by(patient_num, cohort) %>%
  group_modify(count_sequences_hospitalisation)

clinical_raw <- left_join(clinical_raw,
                          hospitalisations_seq_df,
                          by = c("patient_num", "cohort", "days_since_admission"))
```

### Read the ICD10 codes file
The file containing the ICD10 psychiatric related codes is located in the `public-data` folder of the GitHub repository. 

```{r}
#icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codes.csv", header = TRUE, colClasses = "character") %>% 
#  rename(disorder_group = Mental.Health.Disorder.Group, 
#         description = Description, 
#         originalCode = ICD10_Code) %>%
#  mutate( ICD10_Code= ifelse( nchar( originalCode) > 3, paste0(substr( originalCode, 1, 3), ".", substr(  originalCode, 4, nchar(originalCode))),originalCode)) %>%
#  select( disorder_group, ICD10_Code, description)

icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
   select( disorder_group, ICD10_Code, description )
```

## All cohort
We are selecting encounters based on the following criteria:
- associated with at least 1 ICD codes
- with a duration of hospitalization spanning over at least two consecutive days
- Starting on or after 2019/01/01

```{r}
# Adjusting patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
last_data_date <- max(as.Date(demo_raw$last_discharge_date, "%Y-%m-%d"))
clinical_raw <- left_join(clinical_raw, demo_raw[c("patient_num", "cohort", "age", "sex")], by = c("patient_num", "cohort"))
clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)
clinical_raw <- select(clinical_raw, -age)

clinical_raw_filter <- clinical_raw %>%
  filter( len_hospitalisation > min_hosp_days) %>%
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = as.Date(cut( date, breaks = "week")),
                 month = as.Date(cut( date, breaks = "month")),
                 year = format( date, "%Y"), 
                 period = ifelse( date < pandemic_start_date,
                                  "before_pandemic","during_pandemic")) %>% 
  dplyr::filter(cohort  %in% c("AllAdm"))

```

```{r}
all_codes <- clinical_raw_filter %>% 
  left_join( obs_raw,  by = c("patient_num", "cohort", "days_since_admission")) %>%
  filter( concept_type == 'DIAG-ICD10')
stopifnot(unique(all_codes$in_hospital) == 1)
stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
```

## First estimation: total number of patients without filtering by any disease subtype

### New hospitalizations
```{r, fig.width=12, fig.height=12}
new_hospitalizations_per_day <- all_codes %>%
  group_by(patient_num, n_hospitalisation) %>% slice(which.min(date)) %>%
  filter( in_hospital == 1, 
         date < '2021-05-01' &
         date >= '2019-03-01' ) %>%
    group_by( date, weeks, month) %>%
  summarise(count = n_distinct(patient_num)) %>% 
  ungroup()

new_hospitalizations_per_week <- new_hospitalizations_per_day %>%
  filter(weeks >= '2019-03-01') %>% 
  group_by( weeks) %>%
  summarise(count = sum(count))


new_hospitalizations_per_month <- new_hospitalizations_per_day %>%
  group_by( month) %>%
  summarise(count = sum(count))


nh_day <- ggplot(new_hospitalizations_per_day,
                 aes(x = date, y = count))  +
  geom_point() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "New admissions (per day)")

nh_week <- ggplot(new_hospitalizations_per_week, aes(x = weeks,
                       y = count))  +
  geom_point() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "New admissions (per week)")

nh_month <- ggplot(new_hospitalizations_per_month, aes(x = month,
                       y = count))  +
  geom_point() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "New admissions (per month)")

require(gridExtra)
grid.arrange(nh_day,
             nh_week,
             nh_month,
             ncol=1)
```


### Total hospitalize patients
```{r, fig.height=12, fig.width=12}
total_hospitalizations_per_day <- all_codes %>%
  filter(date < '2021-05-01' &
         date >= '2019-03-01' ) %>%
    group_by( date ) %>%
  summarise(count = n_distinct(patient_num))

total_hospitalizations_per_week <- all_codes %>%
  filter(date < '2021-05-01' &
         date >= '2019-03-01' ) %>%
    group_by( weeks ) %>%
  summarise(count = n_distinct(patient_num))

total_hospitalizations_per_month <- all_codes %>%
  filter(date < '2021-05-01' &
         date >= '2019-03-01' ) %>%
    group_by( month ) %>%
  summarise(count = n_distinct(patient_num))



th_day <- ggplot(total_hospitalizations_per_day, aes(x = date,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Total admissions (per day)")

th_week <- ggplot(total_hospitalizations_per_week, aes(x = weeks,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Total admissions (per week)")

th_month <- ggplot(total_hospitalizations_per_month, aes(x = month,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Currently hospitalized patients (per month)")

require(gridExtra)
grid.arrange(th_day, th_week, th_month, ncol=1)
```

## Second estimation: total number of patients filtering by psychiatric disorders
## Creating the inputTable
Creation of the inputTable that will be the only table used throughout the rest of the analysis
- selecting only visits with at least one ICD codes pertaining to the mental health categories
- filter visits where patients are between 11 and 17 years old
- add the demographic and hospitalization related information 
- add the ICD code description and the related disorder group


### Check that we have patients with the ICD10 mental codes
```{r}
### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
codesToReview <- obs_raw %>% dplyr::filter(concept_type == "DIAG-ICD10",
                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", concept_code))
print(codesToReview)

### Calculate how many of the mental codes are in the dataset
mentalHealthCodes <- unique( all_codes %>%
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 concept_code %in% icdCodes$ICD10_Code ) %>%
  dplyr::mutate( concept_code = as.character( concept_code ), 
                 concept_length = nchar( concept_code ), 
                 concept_check = ifelse( concept_length == 4 | concept_length > 8, "Check", "OK")) %>%
  dplyr::select( concept_code, concept_length, concept_check) )

summary(as.factor( mentalHealthCodes$concept_check ) )

print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
```

### Add the mental ICD code description and filter 
```{r message=FALSE, warning=FALSE}
# Getting site name, will be used later in the analysis
site <- unique(as.character(demo_raw$siteid))

icdCategory <- read.delim("public-data/icd10Codes.txt", header = FALSE, colClasses = "character")
colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")

all_codes_desc <- all_codes %>% 
    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>%
  dplyr::select( patient_num, cohort, sex, age_time_visit, concept_type, concept_code, calendar_date,
                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
                 period, disorder_group, description, icd_code_category )

all_codes_desc <- all_codes_desc %>% 
  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))

all_codes_desc <- left_join( all_codes_desc, icdCategory) %>%
  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others")) 

input_data <- all_codes_desc %>% 
  filter(concept_code %in% icdCodes$ICD10_Code)

if( raceAvailable == TRUE ){
  input_data <- input_data %>%
    left_join( race_raw, by = c("patient_num", "cohort") )
}
```

#### Psychiatric related conditions vs. any other condition

##### New admissions


##### Total hospitalizations
We have estimated the total number of hospitalizations in the previous steps:
- total_hospitalization_per_day
- total_hospitalization_per_week
- total_hospitalization_per_month

We will use these counts to estimate the percentage and to compare total counts, vs psychiatric related conditions. 

```{r}
total_hospitalizations_per_day_psy <- input_data %>%
  filter(date < '2021-05-01' &
         date >= '2019-03-01' ) %>%
    group_by( date ) %>%
  summarise(count_psy = n_distinct(patient_num)) %>%
  left_join( total_hospitalizations_per_day ) %>%
  mutate( period = ifelse( date < pandemic_start_date,
                                  "before_pandemic",
                                 "during_pandemic"), 
                percentage = count_psy * 100 / count)

thpd_psyPer <- total_hospitalizations_per_day_psy %>%
  ggplot(aes(x = date, y = percentage, fill = period)) +
  #geom_line()+
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by day)",
       title = "Percentage of patients with mental health related ICD codes (daily)") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

thpd_psyCount <- total_hospitalizations_per_day_psy %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = date, y = value, fill = counts, color = counts)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Date (by day)",
       title = "Daily patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

```

We will compare the psychiatric conditions with some other disorders (e.g, digestive, respiratory, etc.)
We will use these counts to estimate the percentage and to compare total counts, vs psychiatric related conditions. 

```{r}
total_hospitalizations_per_day_disorder_group <- all_codes_desc %>%
  filter(date < '2021-05-01' &
         date >= '2019-03-01' & 
        icd_description_category %in% c("Psy","Neoplams","Diseases of the respiratory system","Diseases of the digestive system" ))%>%
    group_by( date, icd_description_category ) %>%
  summarise(counts = n_distinct(patient_num))

thpd_disorderGroup <- total_hospitalizations_per_day_disorder_group %>%
  ggplot(aes(x = date, y = counts, fill = icd_description_category, color = icd_description_category)) +
  #geom_point(alpha=0.2) +
  geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "patient number",
       x = "days",
       title = "Total number of hospitalized patients (daily)") +   
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1), legend.position = "bottom", legend.text = element_text(size=7), legend.title = element_text(size=7) ) +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
