---
title: "pediatricResults"
author: "aGutierrezSacristan"
date: "10/26/2021"
output: html_document
---
```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}
listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis")
paket(listOfPackages)
```

```{r}
# options(repr.plot.width = 18, repr.plot.height = 9)
theme_plots <- theme_bw() +
  theme(strip.text = element_text(size = 5),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 10),
        plot.subtitle = element_text(size = 9, face = "italic")) 
theme_set(theme_plots)

# Colorblind palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# cut-off dates used in the analysis, these should not be changed
time_period <-"months" #possible values: days, weeks, months
history_start_date <- as.Date("2019-01-01")
history_end_date <- as.Date("2021-05-31")
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2019-01-15")
end_date_plots <- as.Date("2021-05-01")
```


## Read the input file by site
Each site shared the aggregated counts results. Putting all the results together. 

```{r}
outputDir <- "../output/"
outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
getwd()
```



Put all the aggregated counts from the different sites together
```{r}
for( i in 1:length( outputFiles )){
  load(paste0( outputDir, outputFiles[i]))
  if( i == 1 ){
    count_icd_all <- count_icd
    patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc
    patient_count_psy_period_psy_all <- patient_count_psy_period_psy
    perc_disorder_group_all <- perc_disorder_group
    ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy
  }else{
    count_icd_all <- rbind(count_icd_all, count_icd)
    patients_hospitalized_agg_sex_perc_all <- rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc)
    patient_count_psy_period_psy_all <- rbind( patient_count_psy_period_psy_all,patient_count_psy_period_psy)
    perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
    ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
  }
  rm(count_icd,patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
     perc_disorder_group,ratios_patients_with_without_psy,
     patient_count_psy_period_psy_clear, table1, bootstrapped_coefficients_df,
     bootstrapped_fitted_df, length_hospitalisation_values)
}
```


## Patient Count with Mental Health Related ICD codes

```{r}
thpd_psyCount <- count_icd_all %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
  facet_grid(siteid~.) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount
```

#### Both results together

```{r}
count_icd_all_combined <-  count_icd_all %>%
  group_by(time_p, period) %>% 
  summarise(count_psy_all = sum(count_psy), 
            count_icd_all = sum( count_icd ))

thpd_psyCount_combined <- count_icd_all_combined %>%
  pivot_longer(names_to = "counts", cols=c(count_psy_all, count_icd_all)) %>%
  mutate( counts = ifelse( counts == "count_psy_all", "Psy", "Total")) %>%
  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount_combined
```

## Ratio patients with vs without psychiatric conditions
```{r}
ratio_psy_non_psy <- ratios_patients_with_without_psy_all %>%
  ggplot(aes(x = time_p, y = ratio, fill = period, color = period)) +
  geom_point() +
  facet_grid(siteid~.) +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Ratio (with psy conditions/without psy conditions)",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))

ratio_psy_non_psy
```


## Percenge patients with vs without psychiatric conditions
```{r}
percentage_psy_non_psy <- ratios_patients_with_without_psy_all %>%
  ggplot(aes(x = time_p, y = percentage, color = period)) +
  facet_grid(siteid~.) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Percentage",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Percentage patient with vs total population ( per ", time_period,")")) +
  guides(color = guide_legend(title="Patient diagnoses types"))

percentage_psy_non_psy
```

### Test meta-analysis percentage

```{r}
#split this into before and after pandemic and run it twice (one for each subset)

#estimate the summary effect size
ratios_patients_with_without_psy_all$total <- ratios_patients_with_without_psy_all$count_no_psy_patients + ratios_patients_with_without_psy_all$count_psy_patients

ratios_bp <- ratios_patients_with_without_psy_all %>%
  filter( period == "before_pandemic")

ratios_dp <- ratios_patients_with_without_psy_all %>%
  filter( period == "during_pandemic")

metainputdata <- ratios_dp
#RR no transformation
#PLO the logit transformation
#PFT the double arcsine transformation
ies=escalc(xi=count_psy_patients, ni=total, data=metainputdata, measure="PR")
summary(ies$vi)

# pool the individual effect size
# DL random effects using the DerSimonian-Laird estimator
#REML random effects using the restricted maximum-likelihood estimator
pes = rma(yi, vi, data=ies, method="REML")
print(pes)
confint(pes)

pes.summary=metaprop(count_psy_patients, total, as.character(time_p),
                     data=metainputdata, 
                     sm="PRAW")
forest(pes.summary,layout = "JAMA")
```

### Boxplot showing ratio per month

```{r}
ratios_patients_with_without_psy_all %>%
  ggplot( aes(x=as.factor(time_p), y=ratio)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=0.4, alpha=0.9,
                aes(shape = siteid)) +
    theme(
      legend.position="bottom",
      plot.title = element_text(size=11)
    ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
    ggtitle("Ratio boxplot") +
    xlab("")
```

#### Overlap lines: one per site
```{r}
ratios_patients_with_without_psy_all %>%
  ggplot( aes(x=as.factor(time_p), y=ratio, 
              group = siteid, color = siteid))+
  geom_point() + 
  geom_line() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    theme(
      legend.position="bottom",
      plot.title = element_text(size=11)
    ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
    ggtitle("Dot and line plot") +
    xlab("period (months)")

```


#### Meta-analysis of temporal trends




```{r}
## Estimate sampling variances
meta_df <- ratios_patients_with_without_psy_all
meta_df$time <- as.numeric(as.factor(meta_df$time_p))
##### Sampling model 
trend_model <- "percentage ~ period * time"
trends <- meta_df %>% 
  group_by(siteid) %>% 
  group_modify( ~ broom::tidy(lm(as.formula(trend_model), data = .x))) %>%
  filter(term == c("periodduring_pandemic:time")) %>% 
  mutate(variance = std.error * std.error) %>% 
  select(siteid, estimate, variance)
  
```


```{r}
meta_df$sampling_variances
## Fixed Effect model

model_meta <- metafor::rma(yi = estimate, 
             vi = variance,
             measure = "GEN",
  data = trends) 

model_meta %>% 
  broom::tidy()

predict(model_meta)
fitted(model_meta)

metafor::forest(model_meta,
                atransf=exp,
                slab = trends$siteid,
                header="Sites",
                mlab="Overall temporal trend")
```
